Undecidability and the structure of the Turing degrees

### 2018-08-15 paper.tex
Degrees of cardinality continuum

\section{Introduction}
Historical overview of the field, motivation for study
\cite{ambos-spies06:_degrees_unsol}
\cite{soare16_turin_comput}
\cite{soare1999history}
\cite{lerman16:_degrees_unsol}
\todo{}
\subsection{Decision Problems}

@@

\begin{thm}[Closure Properties]
  \label{dcfl:complement}
  If $L$ is \ac{DCFL}, then $\overline{L}$ is a
  \ac{DCFL}. \todo{}
\end{thm}

@@

\begin{lem}
  The relation $\leq_T$ is a partial ordering of $\mathcal{D}$.
\end{lem}

@@

\subsection{Properties and Structure}
\begin{lem}
  The set $\mathcal{D}$ is of cardinality $2^{\aleph_0}$.
\end{lem}

\begin{lem}
  Every Turing degree contains countably infinite elements.
\end{lem}

>>>>

\section{Introduction}
Historical overview of the field, motivation for study
\cite{ambos-spies06:_degrees_unsol}
\cite{soare16_turin_comput}
\cite{soare1999history}
\cite{lerman16:_degrees_unsol}
\todo{}
\subsection{Decision Problems}
What does it mean to \emph{solve a problem}? What is a \emph{problem}, anyways?
These questions may seem superfluous, they must be addressed in order to build a
well defined theory of the difficulty of problems. In this paper, we discuss a
specific type of problem, the \emph{decision problem}. Decision problems are
essentially sets of natural numbers, and to solve the problem is to be able to
decide whether any given element is inside the set or not. Because set theory is
commonly used as the foundational system of mathematics, we can convert problems
expressed in terms of more complex mathematical objects into decision problems,
and thus by studying decision problems can build a system that is able to assess
the computability of all mathematical problems. For example, we can express
strings over the alphabet $\{0,1\}$ as sets of natural numbers, and thus can
study problems about languages of such strings with the same tools we use for
decision problems. Going further, we can express the source code of a computer
program as a string over the alphabet $\{0,1\}$, and so can consider questions
about the behavior of computer programs.

@@

\begin{thm}[Closure Properties]
  \label{dcfl:complement}
  If $L$ is \ac{DCFL}, then $\overline{L}$ is a
  \ac{DCFL}.
  \begin{proof}
    Let $L$ be a \ac{DCFL}. Then, there exists some \ac{DPDA} $P$ that
    recognizes $L$. We can construct a new \ac{DPDA} $P'$ by complementing the
    accepting states of $P$. Because $P$ is a \ac{DPDA} and thus is only in one
    state at a time, $P'$ will recognize $\overline{L}$ and thus $\overline{L}$
    is a \ac{DCFL}. This would not hold if $P$ were only a \ac{PDA}, since a
    nondeterministic \ac{PDA} accepts if \emph{any} branch ends on an accepting
    state, and thus complementing the accepting states would not recognize the
    complement of the originally recognized language.
\end{proof}
\end{thm}

@@

\begin{fact}
  The relation $\leq_T$ is a partial ordering of $\mathcal{D}$.
\end{fact}

@@

\subsection{Properties and Structure}

\begin{lem}
  \label{lem:deg_count}
  Every Turing degree is of cardinality $\aleph_0$.
\end{lem}

\begin{lem}
  The set $\mathcal{D}$ is of cardinality $2^{\aleph_0}$.
  \begin{proof}
    First, observe that the cardinality of $\mathcal{D}$ is at most
    $2^{\aleph_0}$, since it is a set of sets of decision problems and there are
    at most $2^{\aleph_0}$ decision problems since decision problems are subsets
    of $\mathbb{N}$.\par Now, recall from Lemma~\ref{lem:deg_count} that each
    degree contains $\aleph_0$ decision problems. From this, it follows that
    there must be at least $2^{\aleph_0}$ degrees. So, we have that $\lvert \mathcal{D} \rvert=2^{\aleph_0}$.
  \end{proof}
\end{lem}
